---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bracketpredictr

<!-- badges: start -->
[![R-CMD-check](https://github.com/sethbeckett/bracketpredictr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/sethbeckett/bracketpredictr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Introduction

The `` bracketpredictr `` package provides a set of functions for analyzing Division I college basketball stats, from 2013 to 2019. With this package, you can easily retrieve team and game stats for a given season (or multiple), predict winners of upcoming games, and more. **[Here](https://github.com/sethbeckett/bracketpredictr)** is a link to our gitHub repo.

## Installation

To install `bracketpredictr`, you can use the following commands, for installing from github or locally:

```{r, label="install", eval=FALSE}
library(devtools)
install_github("sethbeckett/bracketpredictr")
install.packages("bracketpredictr")
```

## Data - `cbb.csv` (College Basketball Dataset)

This is a combined dataset of Division I basketball teams, across 6 seasons, and 22 variable stats.

## Example usage

Here are some examples of how to use the functions provided by the `bracketpredictr` package:

### Retrieving team stats

To retrieve the names of all available team stats, you can use the `get_stats_names()` function:

```{r}
library(bracketpredictr)
get_stats_names()
```

This will output a list of all available teams, sorted alphabetically.

To retrieve team stats for one or more teams, you can use the `get_team_stats()` function:

```{r}
teams <- c("Duke", "North Carolina", "Kentucky")
stats <- c("TEAM", "W", "ADJOE", "ADJDE")
years <- seq(2013, 2019)

get_team_stats(teams, stats, years)
```

This will output a table showing the requested stats for the specified teams and years. If only wanting one team, stat, or year, you can input just that individual item instead of a vector.

### Predicting game winners

To predict the winner of a game between two teams, you can use the `predict_winner()` function:

```{r}
team1 <- "Duke"
team2 <- "North Carolina"
stats <- c("ADJOE", "ADJDE")
year <- 2019

predict_winner(team1, team2, stats, year)
```

This will output the predicted winner of the game based on the specified stats and year.

## Conclusion

The bracketpredictr package provides a convenient set of functions for analyzing college basketball stats.
