---
title: "bracketpredictr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bracketpredictr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The `` bracketpredictr `` package offers a comprehensive set of functions for analyzing Division I college basketball stats, from 2013 to 2019. The package allows users to easily retrieve team and game stats for a given season or multiple seasons and predict the winners of upcoming games. This can be particularly helpful for those interested in performing in-depth analysis on college basketball performance and making informed predictions for tournaments such as the NCAA March Madness. You can find the Github repository for this package **[here](https://github.com/sethbeckett/bracketpredictr)**[https://github.com/sethbeckett/bracketpredictr].

## Installation

To install `bracketpredictr`, use the following commands for installing from github or locally:

```{r, label="install", eval=FALSE}
library(devtools)
install_github("sethbeckett/bracketpredictr")
install.packages("bracketpredictr")
```

## Data - `cbb.csv` (College Basketball Dataset)

This is a combined dataset of Division I basketball teams, across 6 seasons, and 22 variable stats (such as ORB, ADJOE, and SEED). It includes the years 2013-2019, with 2020 specifically not included in the dataset because of a lack of a postseason due to COVID-19. 2013-2019 was manipulate (cleaned) with the intent of just those years being analyzed together, so for the sake of consistence, the year 2021 was omitted from being add. 

### Source of Data
*Andrew Sundberg. (3/15/2021). College Basketball Dataset, Version 4. Retrieved 3/31/2023 from https://www.kaggle.com/datasets/andrewsundberg/college-basketball-dataset*

## Example usage
Here are some examples of how to use the functions provided by the `bracketpredictr` package:

### Retrieving team stats

To retrieve the names of all available team stats, you can use the `get_stats_names()` function:

```{r}
library(bracketpredictr)
get_stats_names()
```

This will output a list of all available teams, sorted alphabetically.

To retrieve team stats for one or more teams, use the `get_team_stats()` function:

```{r}
teams <- c("Duke", "North Carolina", "Kentucky")
stats <- c("TEAM", "W", "ADJOE", "ADJDE")
years <- seq(2013, 2019)

get_team_stats(teams, stats, years)
```

This will output a table showing the requested stats for the specified teams and years. If only wanting one team, stat, or year, you can input just that individual item instead of a vector.

### Predicting game winners

To predict the winner of a game between two teams, use the `predict_winner()` function:

```{r}
team1 <- "Utah"
team2 <- "BYU"
stats <- c("ADJOE", "ADJDE")
year <- 2019

predict_winner(team1, team2, stats, year)
```

This will output the predicted winner of the game based on the specified team stats, and year.

## Conclusion

The `bracketpredictr` package offers a efficient and convenient set of functions for analyzing college basketball stats. By providing easy access to past data and predictive capabilities, it enables users to conduct in-depth analyses and make well-informed predictions for college basketball tournaments in the future.
